---
title: "moviestats package vignette"
author: "Jacques Serizay"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_document:
        theme: flatly
        highlight: tango
        preserve_yaml: true
        code_folding: show
        df_print: tibble
        toc_float: true
        toc_collapsed: true
        number_sections: false
        thumbnails: true
        gallery: true
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

# Introduction to moviestats

```{r "chunk-options_genes-classes"}
    knitr::opts_chunk$set(
        echo = TRUE, 
        tidy = 'styler', 
        fig.align = 'center', 
        cache = TRUE, 
        cache.path = '.cache/', 
        include = FALSE
    )
```

This package leverages [The Numbers](https://www.the-numbers.com/) database 
to obtain information about movies. 
It can be used to e xtract a lot of useful numbers such as box-office successes 
(or failures!) and contains functions to plot some results for 
individual movies or to compare multiple movies to each other. 

Firstly, we need to get a list of all available movies in the database. We 
can scrape their movies page using the `availableTitles()` function, 
indicating how many pages we want to scrape (each page contains 100 movies). 
Importantly, the pages here are ranked by the movies budgets. If you are 
looking for some low-budget indie movies, you might be facing some issues...  
  
The `availableTitles()` function return an object of moviesDB class, which is a 
list of length 2, containing a list of available titles in the first element, 
and the associated URLs in the second element.

```{r}
    ## Required libraries
    require(magrittr)
    require(rvest)
    ## Get the 10.000 first movies in The Numbers database
    top100.titles <- availableTitles(n.page = 1)
    head(top100.titles$movies.titles)
    head(top100.titles$movies.urls)
```

For convenience, a list of 5200 available movies has already been compiled 
and is accessible in the `moviesDB` object. This has been done in May 
2019, so bare in mind that it will not contain information on more recent movies.

```{r}
    length(unique(moviesDB$movies.urls))
```

One can easily lookup some titles using the `searchTitle()` function. The function
only returns a proper result if it finds a unique movie matching the input pattern, 
otherwise it just prints the matched titles. For ambiguous cases as depicted below, 
the handy `exact.match` logical argument can be used. 

```{r}
    # Search for a title
    searchTitle(pattern = 'expendables', moviesDB = moviesDB)
    searchTitle(pattern = 'The Expendables', moviesDB = moviesDB)
    searchTitle(pattern = 'The Expendables', moviesDB = moviesDB, exact.match = TRUE)
```

Once the **exact** title of a movie is known, information can be retrieved. Note 
that the **exact** title must be entered, otherwise the `fetchMovie()` function 
will fail. 

```{r}
    # With the wrong title
    theexpendables <- fetchMovie(title = 'The expendables', moviesDB = moviesDB)
    # With the right title - Note the case sensitivity
    theexpendables <- fetchMovie(title = 'The Expendables', moviesDB = moviesDB)
    theexpendables
    class(theexpendables)
    theexpendables %>% str
```

As you can see, the output of the `fetchMovie()` is a `movieMetrics` object. 
Just entering the name of the movieMetrics in R console will output a brief 
summary of the data associated with the movie, and will show the first five 
days of box-office results. The `metrics()` function allows to get the full 
box-office results. 

```{r}
    metrics(theexpendables) %>% head
    identical(metrics(theexpendables), theexpendables$metrics)
```

Now, let's scrape some data for a bit more movies.

```{r}
    # Quick search to find the exact title of the movie The Greatest Showman
    searchTitle(pattern = 'greatest showman', moviesDB = moviesDB)
    greatestshowman <- fetchMovie(title = 'The Greatest Showman', moviesDB = moviesDB)
    
    searchTitle(pattern = 'The Post', moviesDB = moviesDB)
    thepost <- fetchMovie(title = 'The Post', moviesDB = moviesDB)
```

But what if you want to obtain data for a list of movies? Let's say, you want
to get data for all the Star Wars movies? First, you can save a list of all 
the movies matching the input pattern in the `searchTitle` function using the 
`return.list = T` logical argument (FALSE by default). Then the `fetchMoviesList`
can take the list as 

```{r}
    starwars.movies <- searchTitle('star wars', moviesDB = moviesDB, return.list = TRUE)
    starwarsMoviesList <- fetchMoviesList(titles = starwars.movies, moviesDB = moviesDB)
    starwarsMoviesList
    class(starwarsMoviesList)
```