---
title: "moviestats package vignette"
author: "Jacques Serizay"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_document:
        theme: flatly
        highlight: tango
        preserve_yaml: true
        code_folding: show
        df_print: tibble
        toc_float: true
        toc_collapsed: true
        number_sections: false
        thumbnails: true
        gallery: true
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

# Introduction to moviestats

```{r "chunk-options_genes-classes"}
    knitr::opts_chunk$set(
        echo = TRUE, 
        tidy = 'styler', 
        fig.align = 'center', 
        cache = TRUE, 
        cache.path = '.cache/', 
        include = FALSE
    )
```

This package leverages [The Numbers](https://www.the-numbers.com/) database 
to obtain information about movies. 
It can be used to e xtract a lot of useful numbers such as box-office successes 
(or failures!) and contains functions to plot some results for 
individual movies or to compare multiple movies to each other. 

## Building a database

Firstly, we need to get a list of all available movies in the database. We 
can scrape their movies page using the `availableTitles()` function, 
indicating how many pages we want to scrape (each page contains 100 movies). 
Importantly, the pages here are ranked by the movies budgets. If you are 
looking for some low-budget indie movies, you might be facing some issues...  
  
The `availableTitles()` function return an object of moviesDB class, which is a 
list of length 2, containing a list of available titles in the first element, 
and the associated URLs in the second element.

```{r}
    ## Required libraries
    require(magrittr)
    require(rvest)
    ## Get the 10.000 first movies in The Numbers database
    top100.titles <- availableTitles(n.page = 1)
    head(top100.titles$movies.titles)
    head(top100.titles$movies.urls)
```

For convenience, a list of 5200 available movies has already been compiled 
and is accessible in the `moviesDB` object. This has been done in May 
2019, so bare in mind that it will not contain information on more recent movies.

```{r}
    moviesDB <- availableTitles(n.page = 52)
    # These URLs are not formatted the same way than the others...
    faulty.urls <- c(
        "https://www.the-numbers.com/movie/Waterworld#tab=box-office",
        "https://www.the-numbers.com/movie/Asterix-aux-Jeux-Olympiques#tab=box-office", 
        "https://www.the-numbers.com/movie/Dantes-Peak#tab=box-office",
        "https://www.the-numbers.com/movie/Army-of-the-Dead-(TBA-ID)#tab=box-office", 
        "https://www.the-numbers.com/movie/Asterix-et-Obelix-Au-service-de-Sa-Majeste#tab=box-office",
        "https://www.the-numbers.com/movie/Renegades-(2017)#tab=box-office", 
        "https://www.the-numbers.com/movie/Ridiculous-6-The#tab=box-office"
    )
    moviesDB <- moviesDB[-match(faulty.urls, moviesDB$movies.url)]
```

One can easily lookup some titles using the `searchTitle()` function. The function
only returns a proper result if it finds a unique movie matching the input pattern, 
otherwise it just prints the matched titles. For ambiguous cases as depicted below, 
the handy `exact.match` logical argument can be used. 

```{r}
    # Search for a title
    searchTitle(pattern = 'expendables', moviesDB = moviesDB)
    searchTitle(pattern = 'The Expendables', moviesDB = moviesDB)
    searchTitle(pattern = 'The Expendables', moviesDB = moviesDB, exact.match = TRUE)
```

## Fetching data about a movie

Once the **exact** title of a movie is known, information can be retrieved. Note 
that the **exact** title must be entered, otherwise the `fetchMovie()` function 
will fail. 

```{r}
    # With the wrong title
    theexpendables <- fetchMovie(title = 'The expendables', moviesDB = moviesDB)
    # With the right title - Note the case sensitivity
    theexpendables <- fetchMovie(title = 'The Expendables', moviesDB = moviesDB)
    theexpendables
    class(theexpendables)
    str(theexpendables)
```

As you can see, the output of the `fetchMovie()` is a `movieMetrics` object. 
Just entering the name of the movieMetrics in R console will output a brief 
summary of the data associated with the movie, and will show the first five 
days of box-office results. The `metrics()` function allows to get the full 
box-office results. 

```{r}
    metrics(theexpendables) %>% head
    identical(metrics(theexpendables), theexpendables$metrics)
```

## Adapted functions

Several functions methods have been adapted for the movieMetrics class. For 
instance, one can subset a movieMetrics object as follows:

```{r}
    # Get the first 2 days
    theexpendables[1:2]
    # Bit more complex: filter by date
    theexpendables %>% 
        filter.movieMetrics(as.Date(c("2010-09-01", "2010-09-30")))
```

One can also get the number of days recorded in the movieMetrics boxoffice
using the `length()` function: 

```{r}
    theexpendables %>% 
        length()
    theexpendables %>% 
        filter(as.Date(c("2010-09-01", "2010-09-30"))) %>% 
        length()
```

## Plotting options

```{r}
    require(ggplot2)
    # Let's compare  The Expendables (2010) and The Greatest Showman (2017)
    theexpendables <- fetchMovie(title = 'The Expendables', moviesDB = moviesDB)
    greatestshowman <- fetchMovie(title = 'The Greatest Showman', moviesDB = moviesDB)
    p1 <- plot(theexpendables, plotly = FALSE)
    p2 <- plot(greatestshowman, plotly = FALSE)
    dev.new(width = 10, height = 5)
    gridExtra::grid.arrange(p1, p2, ncol = 2) # To plot graphs side-by-side
    
    # Wow. looks like The Expendables had a better start. 
    # But let's scale the x axis to the same time window. 
    p1 <- plot(theexpendables, plotly = FALSE, time.range = length(greatestshowman)) + ylim(c(0, 10))
    p2 <- plot(greatestshowman, plotly = FALSE, time.range = length(greatestshowman)) + ylim(c(0, 10))
    gridExtra::grid.arrange(p1, p2, ncol = 2)
    # Now we can see that The Greatest Showman, despite starting lower, stayed in the box-office for much longer!
    
    # What about if we plot the total gross? 
    p1 <- plot(theexpendables, yaxis = 'domestic_total_gross', plotly = FALSE, time.range = length(greatestshowman)) + ylim(c(0, 200))
    p2 <- plot(greatestshowman, yaxis = 'domestic_total_gross', plotly = FALSE, time.range = length(greatestshowman)) + ylim(c(0, 200))
    gridExtra::grid.arrange(p1, p2, ncol = 2)
    
    # And we can also plot the rank in box-office over 90 days from the release date
    p1 <- plot(theexpendables, yaxis = 'rank', plotly = FALSE, time.range = 90) + ylim(c(0, 20))
    p2 <- plot(greatestshowman, yaxis = 'rank', plotly = FALSE, time.range = 90) + ylim(c(0, 20))
    gridExtra::grid.arrange(p1, p2, ncol = 2)
```

## movieMetricsLongList class

We can easily expend our analysis to lists of movies. To fetch data from lots
of movies (> 100), a decent amount of time is needed. We provide an already 
generated list of ~1000 movies with the highest budget in the `top1000.moviesList`
`movieMetricsLongList` object. This list has been generated using the following code.
The `url` argument was used to override the `titles` argument, to be sure to 
retrieve the actual movies urls despite possible doublons in title names. 

```{r}
    movieMetricsLongList <- fetchMoviesList(moviesDB$movies.titles[1:1000], moviesDB)
    str(movieMetricsLongList)
```

```{r}
    MCU.movies <- c(
        "Iron Man", 
        "The Incredible Hulk", 
        "Iron Man 2", 
        "Thor", 
        "Captain America: The First Avenger", 
        "The Avengers (2012)",
        "Iron Man 3", 
        "Thor: The Dark World", 
        "Captain America: The Winter Soldier", 
        "Guardians of the Galaxy", 
        "Avengers: Age of Ultron", 
        "Ant-Man", 
        "Captain America: Civil War", 
        "Doctor Strange", 
        "Guardians of the Galaxy Vol 2", 
        "Spider-Man: Homecoming", 
        "Thor: Ragnarok", 
        "Black Panther",
        "Avengers: Infinity War", 
        "Ant-Man and the Wasp", 
        "Captain Marvel"
    )
    # Fetch data
    marvelLongList <- fetchMoviesList(MCU.movies, moviesDB = moviesDB) 
    # Add MCU phase
    MCU.phase <- factor(c(rep(1, 6), rep(2, 6), rep(3, 9)))
    marvelLongList %<>% addFeature(MCU.phase, "phase")
    # Add year of release
    years <- infos(marvelLongList)$release_date %>% substr(start = 1, stop = 4) %>% as.numeric() %>% factor(., levels = 2008:2019)
    marvelLongList %<>% addFeature(years, "year_of_release")
    # Plot results
    p <- plot.movieMetricsLongList(marvelLongList, x = "day", y = "rank", facet = "year_of_release", xlim = c(0, 120))
```